---
title: Illustration of how to translate dogfish TAC between different discard mortality assumptions
format: pdf
date: "2025-02-11"
---

This document works through how to translate dogfish TAC between different discard mortality assumptions. It is intended as a thought experiment and is not formal advice.

First, we need a new term for this document: "IFMP dead catch". So, we have

- "catch" (all landings, all discards dead or alive)
- "dead catch" (all landings, all discards assumed to have died according to literature-based rates)
- "IFMP dead catch" (all landings, all discards multiplied by the IFMP discard-mortality rate); this is the quantity in the TAC

We also have 

- "science discard mortality rate" a discard rate derived from the scientific literature
- "IFMP discard mortality rate" a discard rate from the IFMP

For this illustration, we assume that we want to set *IFMP dead catch* such that *dead catch* works out to the desired quantity.

```{r, eval=FALSE, include=FALSE}
#| eval: FALSE
#| include: FALSE

library(dplyr)
library(ggplot2)
d <- readRDS("../gfsynopsis-2023/report/data-cache-2024-05/cpue-index-dat.rds")
dd <- filter(d, species_common_name == "SPINY DOGFISH")
dd <- mutate(dd,
      hours_fished =
        as.numeric(difftime(fe_end_date, fe_start_date, units = "hours"))
    )
median(dd$hours_fished, na.rm = TRUE)
mean(dd$hours_fished, na.rm = TRUE)
quantile(dd$hours_fished, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
quantile(dd$hours_fished, probs = c(0.1, 0.5, 0.9), na.rm = TRUE)

more_than3 <- sum(dd$hours_fished > 3, na.rm = TRUE)
all_tows <- nrow(dd)
more_than3/all_tows
```

The IFMP has the following discard mortality assumptions:

Longline:

- 6% discard mortality for longline gear

Trawl:

- 5% mortality for the first 2 hours of a trawl tow
- an additional 5% mortality for each subsequent hour

For all trawl tows that have caught dogfish, the median tow length is 1.8 hours and the 25th and 75th quantiles are 1.1 and 2.5 hours.
Let's keep it simple for trawl and call it 7.5% since ~50% of trawls are under 2 hours and ~50% are more. 
Only ~14% of tows are greater than 3 hours.
So, if a trawl event catches 1000 t and discards it all, that gets recorded as 1000 t * 0.075 = 75 t of IFMP dead dogfish against quota.
If, instead, we use a discard mortality rate of 37% within the model for a given gear type, we're assuming that 1000 t of catch actually corresponds to 1000 t * 0.37 = 370 t of dead catch.

So, if GMU wanted to achieve a maximum of 370 t dead catch, GMU would have to set IFMP dead catch (i.e., TAC) to 370 * (7.5 / 37) = 75 t to achieve the desired dead catch.

In conclusion, the formula is:

```
IFMP_dead_catch = desired_dead_catch * IFMP_discard_rate / science_discard_rate
```

For longline, the assessment uses 8%, 27%, and 36% for longline and 19%, 37%, and 56% for trawl.

If desired maximum dead catch was 200 t across all fleets and we wanted to allocate 50% of that to the longline fleet (a completely arbitrary value here), let's see how that 100 t would look under low, base, and high *science discard mortality rates*:

IFMP dead catch for longline:

```{r}
round(100 * 6 / 8)
round(100 * 6 / 27)
round(100 * 6 / 36)
```

IFMP dead catch for trawl:

```{r}
round(100 * 7.5 / 19)
round(100 * 7.5 / 37)
round(100 * 7.5 / 56)
```
