---
title: "Dogfish catch comparisons"
execute:
  echo: false
  message: false
  warning: false
  results: 'hide'
  fig.asp: 1.618
format: 
  html:
    self-contained: true
---

```{r}
#| echo: false
```

```{r setup}
library(dplyr)
library(ggplot2)
library(readxl)
theme_set(theme_bw())
library(gfplot)
library(here)
```

# Trawl data

Plot of raw data as sent by data unit:

```{r, fig.asp=0.5}
d <- read_xlsx(here("data/raw/FD5046 dogfish trawl catch 1996-2024.xlsx"), sheet = 1L)
names(d) <- tolower(names(d))
dsum <- group_by(d, gear_subtype, calendar_year) |>
  summarise(landed = sum(landed_round_kg)/1000, released = sum(total_released_round_kg)/1000)

dsum |> tidyr::pivot_longer(cols = c(landed, released), names_to = "type") |>
  ggplot(aes(calendar_year, value, fill = type)) + geom_col() +
  facet_wrap(~gear_subtype) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Catch (t)") + xlab("Calendar year")
```

```{r}
catch <- readRDS(here("data/generated/catch-dec2024.rds")) %>%
  group_by(year, gear) %>%
  summarise(landing = 1e-3 * sum(landed_kg),
    discard = 1e-3 * sum(discarded_kg), .groups = "drop") %>%
  ungroup()
bottom_trawl_gear <- c("Trawl + hook and line", "Trawl", "Bottom trawl", "Unknown/trawl")
f1 <- catch %>%
  filter(gear %in% bottom_trawl_gear) %>%
  select(year, landing, discard) %>%
  summarise(value = sum(landing), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 1, type = "landings", fleet_name = "Bottom trawl")
f2 <- catch %>%
  filter(gear %in% bottom_trawl_gear) %>%
  select(year, landing, discard) %>%
  summarise(value = sum(discard), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 2, type = "discards", fleet_name = "Bottom trawl")
# Due to initial fleet set-up, discards + landings were combined for midwater trawl
# Incorporate alternative discard mortality values into the total catch series
# For all other fleets, incorporate the discard mortality in the SS3 control file
# disc_mort <- midwater_discard_rate
disc_mort <- 1
f3 <- catch %>%
  filter(gear == "Midwater trawl") %>%
  select(year, landing, discard) %>%
  summarise(value = sum(landing), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 3, type = "landings", fleet_name = "Midwater trawl")
f4 <- catch %>%
  filter(gear == "Midwater trawl") %>%
  select(year, landing, discard) %>%
  summarise(value = sum(discard), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 3, type = "discards", fleet_name = "Midwater trawl")
out <- rbind(f1, f2, f3, f4) %>%
  mutate(value = round(value, 4)) %>%
  # mutate(se = 0.01,
  #   season = 1) %>%
  filter(value > 0) |>
  select(year, fleet_name, type, value) %>%
  arrange(fleet_name, year) |>
  mutate(source = "RPR meeting")

out2 <- dsum |> mutate(source = "Data unit")
out2$gear_subtype[out2$gear_subtype == "BOTTOM TRAWL"] <- "Bottom trawl"
out2$gear_subtype[out2$gear_subtype == "UNSPECIFIED"] <- "Bottom trawl"
out2$gear_subtype[out2$gear_subtype == "MIDWATER TRAWL"] <- "Midwater trawl"
out2 <- group_by(out2, calendar_year, source, gear_subtype) |>
  summarise(landed = sum(landed), released = sum(released), .groups = "drop")
out2 <- rename(out2, year = calendar_year, fleet_name = gear_subtype, discards = released, landings = landed)
out2 <- out2 |> tidyr::pivot_longer(cols = c(landings, discards), names_to = "type")
combined <- bind_rows(out, out2)
```

Comparison between data unit and RPR trawl data:

```{r}
combined |>
  filter(year >= 1996, year <= 2023) |>
  mutate(type = forcats::fct_rev(type)) |>
  ggplot(aes(year, value, fill = type)) + geom_col() +
  facet_grid(source~fleet_name) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Catch (t)") + xlab("Calendar year") +
  scale_fill_brewer(palette = "Set2") + labs(fill = "Catch type") +
  scale_x_continuous(breaks = seq(2000, 2025, 5))
```

Bottom trawl look nearly identical. Confirm that:

```{r, fig.asp=0.6}
d <- readRDS(here("data/generated/catch-dec2024.rds"))
d <- filter(d, year >= 1996, year <= 2023)
d <- filter(d, gear %in% c("Bottom trawl"))

# d$area <- gfplot::assign_areas(d$major_stat_area_name, c("5[ABCDE]+", "3[CD]+"))
dd <- group_by(d, year) |> 
  summarise(landings = sum(landed_kg/1000, na.rm = TRUE), discards = sum(discarded_kg/1000))
dd2 <- tidyr::pivot_longer(dd, c("landings", "discards"), names_to = "type") |> 
  mutate(source = "RPR")

bind_rows(
  select(dd2, year, value, source, type),
  select(filter(out2, fleet_name == "Bottom trawl"), year, value, source, type)) |> 
  filter(year >= 1996, year <= 2023) -> combined2

combined2 |> 
  ggplot(aes(year, value, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Catch (t)") + ggtitle("Bottom trawl") + xlab("Year") +
  facet_wrap(~type) + labs(colour = "Source")
ggsave(here("figs/bottom-trawl-gdu.png"), width = 7, height = 4)
```

```{r, fig.asp=0.6}
d <- readRDS(here("data/raw/catch-2024-12-09.rds"))
d <- filter(d, year >= 1996, year <= 2023)
d <- filter(d, !grepl("^4B", major_stat_area_name))
d <- filter(d, gear %in% c("BOTTOM TRAWL", "UNKNOWN TRAWL"))

# d$area <- gfplot::assign_areas(d$major_stat_area_name, c("5[ABCDE]+", "3[CD]+"))
dd <- group_by(d, year) |> 
  summarise(landings = sum(landed_kg/1000, na.rm = TRUE), discards = sum(discarded_kg/1000))
dd2 <- tidyr::pivot_longer(dd, c("landings", "discards"), names_to = "type") |> 
  mutate(source = "RPR + all non-4B")

bind_rows(
  select(dd2, year, value, source, type),
  select(filter(out2, fleet_name == "Bottom trawl"), year, value, source, type)) |> 
  filter(year >= 1996, year <= 2023) -> combined2

combined2 |> 
  ggplot(aes(year, value, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Catch (t)") + ggtitle("Bottom trawl") + xlab("Year") +
  facet_wrap(~type) + labs(colour = "Source")
ggsave(here("figs/bottom-trawl-gdu2.png"), width = 7, height = 4)
```

Midwater trawl landings look different pre 2007. Make a simpler data pipeline from the merged_catch table for comparison:

```{r, fig.asp=0.6}
d <- readRDS(here("data/raw/catch-2024-12-09.rds"))
d <- filter(d, !grepl("^4B", major_stat_area_name))
d <- filter(d, year >= 1996, year <= 2023)
d <- filter(d, gear %in% c("MIDWATER TRAWL"))
d <- filter(d, dfo_stat_area_code != 12)
d$area <- gfplot::assign_areas(d$major_stat_area_name, c("5[ABCDE]+", "3[CD]+"))
dd <- group_by(d, year) |> 
  summarise(landings = sum(landed_kg/1000, na.rm = TRUE), discards = sum(discarded_kg/1000))
dd2 <- tidyr::pivot_longer(dd, c("landings", "discards"), names_to = "type") |> 
  mutate(source = "RPR + all non-4B")

bind_rows(
  select(dd2, year, value, source, type),
  select(filter(out2, fleet_name == "Midwater trawl"), year, value, source, type)) |> 
  filter(year >= 1996, year <= 2023) -> combined2

combined2 |> 
  ggplot(aes(year, value, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Landings (t)") + ggtitle("Midwater trawl") + xlab("Year") +
  facet_wrap(~type)
```

Try again but this time add in the GFBio midwater trawl 'total_kg' since it's not split into landings/discards:

```{r}
d <- read_xlsx(here("data/raw/FD5046 dogfish trawl catch 1996-2024.xlsx"), sheet = 1L)
names(d) <- tolower(names(d))

d <- mutate(d, landed_round_kg = if_else(database_name == "GFBio" & gear_subtype == "MIDWATER TRAWL" & total_kg > 0 & landed_round_kg == 0 & total_released_round_kg == 0, total_kg, landed_round_kg))
dsum <- group_by(d, gear_subtype, calendar_year) |>
  summarise(landed = sum(landed_round_kg)/1000, released = sum(total_released_round_kg)/1000)
out3 <- dsum |> mutate(source = "Data unit")
out3$gear_subtype[out3$gear_subtype == "BOTTOM TRAWL"] <- "Bottom trawl"
out3$gear_subtype[out3$gear_subtype == "UNSPECIFIED"] <- "Bottom trawl"
out3$gear_subtype[out3$gear_subtype == "MIDWATER TRAWL"] <- "Midwater trawl"
out3 <- group_by(out3, calendar_year, source, gear_subtype) |>
  summarise(landed = sum(landed), released = sum(released), .groups = "drop")
out3 <- rename(out3, year = calendar_year, fleet_name = gear_subtype, discards = released, landings = landed)
out3 <- out3 |> tidyr::pivot_longer(cols = c(landings, discards), names_to = "type")

bind_rows(
  select(dd2, year, value, source, type),
  select(filter(out3, fleet_name == "Midwater trawl"), year, value, source, type)) |> 
  filter(year >= 1996, year <= 2023) -> combined3

combined3 |> 
  ggplot(aes(year, value, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Landings (t)") + ggtitle("Midwater trawl") + xlab("Year") +
  facet_wrap(~type) + labs(colour = "Source")
ggsave(here("figs/mid-trawl-gdu2.png"), width = 7, height = 4)
```

# Longline data

```{r}
# d1 <- read_xlsx(here("data/raw/FD5046 dogfish line catch 2001-2024.xlsx"), sheet = 1L, na = "NULL")
d1 <- readRDS("~/Downloads/dogfish-ll-with-12.rds")
# names(d1) <- tolower(names(d1))
names(d1) <- tolower(names(d1))

catch <- readRDS(here("data/generated/catch.rds")) %>%
  filter(area != "4B", year <= 2023) %>%
  group_by(year, gear) %>%
  summarise(landing = 1e-3 * sum(landed_kg),
    discard = 1e-3 * sum(discarded_kg), .groups = "drop") %>%
  ungroup()
f4 <- catch %>%
  filter(gear == "Hook and line") %>%
  select(year, landing, discard) %>%
  summarise(value = sum(landing), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 4, type = "landings", fleet_name = "Hook and line")
f5 <- catch %>%
  filter(gear == "Hook and line") %>%
  select(year, landing, discard) %>%
  summarise(value = sum(discard), landing = sum(landing), discard = sum(discard), .by = year) %>%
  mutate(fleet = 5, type = "discards", fleet_name = "Hook and line")

out <- rbind(f4, f5) %>%
  mutate(value = round(value, 4)) %>%
  # mutate(se = 0.01,
  #   season = 1) %>%
  filter(value > 0) |>
  select(year, fleet_name, type, value) %>%
  arrange(fleet_name, year) |>
  mutate(source = "RPR meeting") |> filter(type != "discards")

out2 <- group_by(d1, calendar_year) |>
  summarize(landings = sum(landed_round_kg, na.rm = TRUE) / 1000,
    discards_count = sum(total_released_count, na.rm = TRUE)) |>
  mutate(source = "Data unit") |> rename(year = calendar_year)

out2 <- out2 |> tidyr::pivot_longer(cols = c(landings, discards_count), names_to = "type") |>
  mutate(fleet_name = "Hook and line")

combined <- bind_rows(out, out2)
```

Compare data unit and RPR longline catch weight data:

```{r}
combined |>
  filter(year >= 2001) |>
  filter(type == "landings") |>
  # mutate(type = forcats::fct_rev(type)) |>
  ggplot(aes(year, value, fill = type)) + geom_col() +
  facet_grid(source~fleet_name) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Catch (t)") + xlab("Calendar year") +
  scale_fill_brewer(palette = "Set2") + labs(fill = "Catch type") +
  scale_x_continuous(breaks = seq(2000, 2025, 5))
```

Looks the same.

Here are the discard counts from the data unit:

```{r}
combined |>
  filter(year >= 2006, year <= 2023) |>
  filter(type != "landings") |>
  # mutate(type = forcats::fct_rev(type)) |>
  ggplot(aes(year, value/1000, fill = type)) + geom_col() +
  facet_grid(source~fleet_name) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ylab("Released (1000 dogfish)") + xlab("Calendar year") +
  scale_fill_brewer(palette = "Set2") + labs(fill = "Catch type")
```

Divided by sector:

```{r}
group_by(d1, calendar_year, fishery_sector) |>
  summarise(total_released_count = sum(total_released_count, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(total_released_count)) |>
  filter(calendar_year >= 2006, calendar_year <= 2023) |>
  ggplot(aes(calendar_year, total_released_count/1000, fill = fishery_sector)) + geom_col() +
  scale_fill_brewer(palette = "Set2") + labs(fill = "Sector") +
  ylab("Released (1000 dogfish)") + xlab("Calendar year") +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05)))
ggsave(here("figs/sector-discard-counts.png"), width = 8, height = 5)
```

Try multiplying each fish by 9 lbs:

```{r}
group_by(d1, calendar_year, fishery_sector) |>
  summarise(total_released_count = sum(total_released_count, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(total_released_count)) |>
  filter(calendar_year >= 2006, calendar_year <= 2023) |>
  ggplot(aes(calendar_year, total_released_count*4.08233/1000, fill = fishery_sector)) + geom_col() +
  scale_fill_brewer(palette = "Set2") + labs(fill = "Sector") +
  ylab("Released (t)") + xlab("Calendar year") +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05)))
```

What is the avg. kg/fish from available released data?

```{r, fig.asp=0.6}
filter(d1, total_released_count > 0, total_released_round_kg > 0) |>
  mutate(kg_per_fish = total_released_round_kg / total_released_count) |>
  ggplot(aes(calendar_year, kg_per_fish, colour = gear)) + geom_line() +
  facet_wrap(~fishery_sector)
```

Now compare with the RPR release data. First, there was a bug in MERGED_CATCH, so check how this affected things:

```{r, fig.asp=0.6}
d_4b5abcde3cd <- readRDS(here("data/generated/catch-4B5ABCDE3CD-summarized.rds"))
d_4b5abcde3cd$gear[d_4b5abcde3cd$gear == "Trap"] <- "Hook and line" # FIXME?

drpr <- filter(d_4b5abcde3cd, year >= 2000, year <= 2023) |>
  filter(area != "4B") |> 
  mutate(
    discarded_kg =
      ifelse(
        gear != "Hook and line",
        discarded_kg,
        4.08233 * discarded_pcs # assuming 9 lbs avg. per discarded dogfish
      )
  ) |>
  group_by(year, gear) |>
  summarise(total_released_count = sum(discarded_pcs), landed_t = sum(landed_kg/1000), discards_t = sum(discarded_kg/1000), .groups = "drop") |>
  ungroup() |>
  group_by(year) |>
  arrange(gear) |>
  filter(!gear %in% c("Unknown/trawl", "Midwater trawl", "Bottom trawl")) |> ungroup() |>
  rename(calendar_year = year) |> mutate(source = "RPR")
  # mutate(landed_t = round(landed_t, 1), discards_t = round(discards_t, 1))
  # readr::write_csv("~/Downloads/catch-oct18.csv")

gdu <- group_by(d1, calendar_year) |>
  summarise(total_released_count = sum(total_released_count, na.rm = TRUE), .groups = "drop") |>
  mutate(landed_t = total_released_count * 4.08233 / 1000) |>
  mutate(source = "Data unit")

dd <- bind_rows(
  select(drpr, calendar_year, total_released_count, source),
  select(gdu, calendar_year, total_released_count, source)
)

# dd |> filter(calendar_year >= 2006) |>
#   ggplot(aes(calendar_year, total_released_count)) + geom_col() +
#   facet_wrap(~source)

# with updated merged-catch table!
d <- readRDS(here("data/raw/catch-2024-12-09.rds"))
nrow(d)
d <- filter(d, !grepl("^4B", major_stat_area_name))
d <- filter(d, dfo_stat_area_code != 12)
nrow(d)
source(here("analysis/utils.R"))
d_4b5abcde3cd <- tidy_catch_dogfish(d, areas = c("5[ABCDE]+", "3[CD]+"))
d_4b5abcde3cd$area[d_4b5abcde3cd$area == "3CD"] <- "5ABCDE3CD"
d_4b5abcde3cd$area[d_4b5abcde3cd$area == "5ABCDE"] <- "5ABCDE3CD"
d_4b5abcde3cd$gear[d_4b5abcde3cd$gear == "Trap"] <- "Hook and line"
d_4b5abcde3cd <- group_by(d_4b5abcde3cd, year, species_common_name, area, gear) |>
  summarise(
    landed_kg = sum(landed_kg),
    discarded_kg = sum(discarded_kg),
    discarded_pcs = sum(discarded_pcs), .groups = "drop"
  )
drpr2 <- filter(d_4b5abcde3cd, year >= 2000, year <= 2023) |>
  mutate(
    discarded_kg =
      ifelse(
        gear != "Hook and line",
        discarded_kg,
        4.08233 * discarded_pcs # assuming 9 lbs avg. per discarded dogfish
      )
  ) |>
  group_by(year, gear) |>
  summarise(total_released_count = sum(discarded_pcs), landed_t = sum(landed_kg/1000), discards_t = sum(discarded_kg/1000)) |>
  ungroup() |>
  group_by(year) |>
  arrange(gear) |>
  filter(!gear %in% c("Unknown/trawl", "Midwater trawl", "Bottom trawl")) |> ungroup() |>
  rename(calendar_year = year) |> mutate(source = "RPR with updated MERGED_CATCH table")

dd <- bind_rows(
  select(drpr, calendar_year, total_released_count, source),
  select(drpr2, calendar_year, total_released_count, source)
)

dd |> filter(calendar_year >= 2000) |>
  ggplot(aes(calendar_year, total_released_count, colour = source, group = source)) + geom_line()
```

Looks very similar; a slight decrease.

Now, compare data unit with RPR merged catch (with corrected MERGED_CATCH table):

```{r}
dd <- bind_rows(
  select(drpr2, calendar_year, total_released_count, source),
  select(gdu, calendar_year, total_released_count, source)
)
dd |> filter(calendar_year >= 2006, calendar_year <= 2023) |>
  ggplot(aes(calendar_year, total_released_count)) + geom_col() +
  facet_wrap(~source) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05)))
```

Make a simpler data pipline to compare merged catch with the latest data unit version:

```{r, fig.asp=0.6}
d <- readRDS(here("data/raw/catch-2024-12-09.rds"))
d <- filter(d, fishery_sector != "ROCKFISH INSIDE")
d <- filter(d, !grepl("^4B", major_stat_area_name))
d <- filter(d, year >= 2000, year <= 2023)
d <- filter(d, gear %in% c("HOOK AND LINE", "TRAP", "LONGLINE"))
d$area <- gfplot::assign_areas(d$major_stat_area_name, c("5[ABCDE]+", "3[CD]+"))
dd <- group_by(d, year) |> 
  summarise(total_released_count = sum(discarded_pcs, na.rm = TRUE)) |> 
  mutate(source = "MERGED_CATCH") |> 
  rename(calendar_year = year)

bind_rows(
  select(dd, calendar_year, total_released_count, source),
  select(gdu, calendar_year, total_released_count, source)) |> 
  filter(calendar_year >= 2007, calendar_year <= 2023) |> 
  ggplot(aes(calendar_year, total_released_count/1000)) + geom_col() +
  facet_wrap(~source) +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05)))

bind_rows(
  select(dd, calendar_year, total_released_count, source),
  select(gdu, calendar_year, total_released_count, source)) |> 
  filter(calendar_year >= 2005, calendar_year <= 2023) |> 
  ggplot(aes(calendar_year, total_released_count/1000, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Longline + trap")
```

# What about by sector for longline?

```{r}
d1 <- read_xlsx(here("data/raw/FD5046 dogfish line catch 2001-2024.xlsx"), sheet = 1L, na = "NULL")
d1 <- readRDS("~/Downloads/dogfish-ll-with-12.rds")
# names(d1) <- tolower(names(d1))
names(d1) <- tolower(names(d1))

dd_ll <- d |> 
  filter(gear %in% "HOOK AND LINE") |> 
  filter(!grepl("^4B", major_stat_area_name)) |> 
  group_by(year, fishery_sector) |>
  summarise(total_released_count = sum(discarded_pcs, na.rm = TRUE)) |> 
  mutate(source = "MERGED_CATCH") |> 
  rename(calendar_year = year)

gdu_ll <- d1 |> 
  filter(grepl("LINE", gear)) |> 
  group_by(calendar_year, fishery_sector) |>
  summarise(total_released_count = sum(total_released_count, na.rm = TRUE), .groups = "drop") |>
  mutate(source = "Data unit")

bind_rows(
  select(dd_ll, calendar_year, total_released_count, source, fishery_sector),
  select(gdu_ll, calendar_year, total_released_count, source, fishery_sector)) |> 
  filter(calendar_year >= 2000, calendar_year <= 2023) |> 
  ggplot(aes(calendar_year, total_released_count/1000, colour = source)) + geom_line() +
  scale_y_continuous(lim = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Longline + hook and line") +
  facet_wrap(~fishery_sector, scales = "free_y") +
  ylab("Total dogfish released (1000s)")
  # geom_vline(xintercept = 2007)
ggsave(here("figs/longline-gdu-sector.png"), width = 10, height = 6)
```

So, the problem is spiny dogfish sector longline discards.

# The major issue: what is a good average weight for longline discards?

```{r}
ds <- readRDS(here("data/raw/commercial-samples.rds"))
table(ds$gear_desc)

dsll <- filter(ds, gear_desc == "LONGLINE")

filter(ds) |> 
  filter(!is.na(length)) |> 
  filter(!gear_desc %in% c("SHRIMP TRAWL", "UNKNOWN", "UNKNOWN TRAWL")) |> 
  ggplot(aes(factor(year), length)) + 
# geom_point() +
  facet_wrap(~gear_desc) +
  geom_boxplot() +
  scale_x_discrete(breaks = seq(1900, 2024, 5))
```

Convert lengths to weights:

```{r}
# > mm$pars
# $log_a
# [1] -12.32762
# $b
# [1] 2.96391
# > mf$pars
# $log_a
# [1] -13.18455
# $b
# [1] 3.182569
# lm(log(weight) ~ log(length))

x <- filter(ds) |> 
  filter(!is.na(length)) |> 
  filter(sex %in% c(1, 2)) |> 
  filter(!gear_desc %in% c("SHRIMP TRAWL", "UNKNOWN", "UNKNOWN TRAWL")) |> 
  mutate(predicted_weight = 
    ifelse(
      sex == 1, exp(-12.32762 + log(length) * 2.96391),
      exp(-13.18455 + log(length) * 3.182569))) |> 
  select(length, weight, predicted_weight, sex, gear_desc, year, sampling_desc)

plot(x$weight/1000, x$predicted_weight, ylim = c(0, 3), xlim = c(0, 3));abline(0, 1)

x |> 
  ggplot(aes(factor(year), predicted_weight)) + 
# geom_point() +
  facet_wrap(~gear_desc) +
  geom_boxplot() +
  scale_x_discrete(breaks = seq(1900, 2024, 5))

x |> 
  ggplot(aes(year, predicted_weight)) + 
  geom_point(alpha = 0.01) +
  facet_grid(sampling_desc~gear_desc)

xx <- filter(x, gear_desc == "LONGLINE", sampling_desc %in% c("KEEPERS", "UNSORTED"))

xx |> 
  filter(year > 2000) |> 
  mutate(sex = ifelse(sex ==1, "Male", "Female")) |> 
  ggplot(aes(year, predicted_weight)) + 
  geom_point(alpha = 0.01) +
  facet_grid(sampling_desc~sex) +
  geom_smooth(se = FALSE, method = "lm") +
  # geom_vline(xintercept = c(2009), lty = 2) +
  xlab("Year") + ylab("Weight (kg)") +
  scale_x_continuous(breaks = seq(2000, 2022, 2))
ggsave(here("figs/weight-longline-keepers-unsorted.png"), width = 7, height = 5)

xx |> filter(year %in% c(2009)) |> 
  group_by(sex, sampling_desc) |> summarise(mean = mean(predicted_weight)) |> 
  saveRDS("figs/keepers-unsorted-avg-weight-male-female.rds")
```

The unsorteds are recent and presumably discarded after

```{r}
xx |> filter(sampling_desc == "UNSORTED", year > 2000) |>
  group_by(sex) |> 
  summarise(mean_weight = mean(predicted_weight)) |> 
  mutate(mean_weight_lbs = mean_weight * 2.20462)

o <- xx |> filter(sampling_desc == "UNSORTED", year > 2000) |>
  # group_by(sex) |> 
  summarise(mean_weight = mean(predicted_weight), median_weight = median(predicted_weight)) |> 
  mutate(mean_weight_lbs = mean_weight * 2.20462, median_weight_lbs = median_weight * 2.20462)
saveRDS(o, "figs/keepers-unsorted-avg-median-weight-male-female.rds")

xx |> filter(sampling_desc == "UNSORTED", year > 2000) |> 
  mutate(sex_label = ifelse(sex == 1, "Male", "Female")) |> 
  ggplot(aes(factor(year), predicted_weight, colour = sex_label)) + 
  geom_violin(alpha = 0.3) +
  geom_point(alpha = 0.1, position = position_jitter(height = 0, width = 0.1)) +
  # facet_wrap(~sex_label) +
  labs(y = "Weight (kg)", x = "Year", colour = "Sex") +
geom_hline(yintercept = o$mean_weight, lty = 2)
ggsave(here("figs/longline-unsorted-weight.png"), width = 5, height = 3.4)

xx |> filter(sampling_desc == "UNSORTED", year > 2000) |>
  group_by(sex) |> 
  summarise(mean_weight = mean(predicted_weight)) |> 
  mutate(mean_weight_lbs = mean_weight * 2.20462)

xx |> filter(sampling_desc == "UNSORTED", year > 2000) |>
  # group_by(sex) |> 
  summarise(mean_weight = mean(predicted_weight), median_weight = median(predicted_weight)) |> 
  mutate(mean_weight_lbs = mean_weight * 2.20462, median_weight_lbs = median_weight * 2.20462)
```

## Survey weights:

```{r}
ds <- readRDS(here("data/raw/survey-samples.rds")) |> 
  filter(!is.na(survey_abbrev))

means <- group_by(ds, survey_abbrev) |> 
  summarise(median_weight = median(weight/1000, na.rm = T), mean_weight = mean(weight/1000, na.rm = TRUE)) |> 
  filter(!grepl("INS ", survey_abbrev)) |> 
  filter(!grepl(" MSA", survey_abbrev)) |> 
  filter(survey_abbrev != "SYN WCHG")

filter(ds) |> 
  filter(!grepl("INS ", survey_abbrev)) |> 
  filter(!grepl(" MSA", survey_abbrev)) |> 
  filter(survey_abbrev != "SYN WCHG") |> 
  filter(!is.na(weight)) |> 
  ggplot(aes(factor(year), weight/1000)) + 
  facet_wrap(~survey_abbrev) +
  geom_boxplot() +
  geom_hline(yintercept = 4.08233, lty = 3) +
  geom_hline(aes(yintercept = mean_weight), data = means, col = "red") +
  geom_hline(aes(yintercept = median_weight), data = means, col = "blue") +
  scale_x_discrete(breaks = seq(1900, 2024, 2))
  # geom_point(alpha = 0.005)

means |> 
  filter(survey_abbrev != "SYN WCHG") |> 
  mutate(median_weight_lbs = median_weight * 2.20462, mean_weight_lbs = mean_weight * 2.20462)
```

